Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-05-06 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
hs: Gregorio Carullo`s comments

%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
% General:
The detection of gravitational waves produced by coalescing compact binaries
allows us to explore the dynamical, nonlinear regime of general relativity
and to constrain modifications to it.
%
Some of these gravitational-wave events have sufficiently high
signal-to-noise ratio in the merger, allowing us to probe the relaxation of the remnant black
hole to its final, stationary state --- the so-called black-hole ringdown, which is
characterized by a set of quasinormal modes.
% Main question:
Can we use the ringdown to constrain deviations to general relativity,
as predicted by several of its contenders?
% What we do:
Here, we  address this question by using an inspiral-merger-ringdown waveform
model in the effective-one-body formalism, augmented with a parametrization of
the ringdown based on an expansion in the final black hole's spin.
%
We give a prescription on how to include in this waveform model the quasinormal mode frequencies
calculated on a theory-by-theory basis. In particular, we focus on theories that modify
general relativity by higher-order curvature corrections, namely,
Einstein-dilaton-Gauss-Bonnet and dynamical Chern-Simons theories, and cubic- and quartic-order
effective-field-theories of general relativity.
%
% Main conclusion:
We confront the resultant waveform models against the two loudest
gravitational-wave ringdown signals observed so far, notably, GW150914 and GW200129.
%
We find that while the Einstein-dilaton-Gauss-Bonnet theory cannot be
constrained with these events, we can place upper bounds on the
fundamental length-scale of cubic- ($\ell_{\rm cEFT} \leqslant 38.2$~km) and quartic-order
($\ell_{\rm qEFT} \leqslant 51.3$~km) effective-field-theories of general
relativity, and of dynamical Chern-Simons gravity ($\ell_{\rm dCS} \leqslant 38.7$~km).
%
The latter result is a concrete example of a theory presently unconstrained by
inspiral-only analyses which, however, can be constrained by merger-ringdown
studies with current gravitational-wave data.
%
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%

@@

In the end, the $\pSEOB$ waveform model with theory-specific
QNMs has only $\ell_{\rm th}$ as a free beyond-GR parameter.
%
We emphasize that our procedure is different from that of
Ref.~\cite{Carullo:2021dui} which, for a given value of $p$, varied all $\ell_{\rm th}$,
$\delta\omega^{(j)}$, and $\delta\tau^{(j)}$ parameters, and then proceeded
to use the posteriors on $\ell_{\rm th}$ to draw conclusions to any theory that predict
corrections at a given $p$.
%
This is a important difference between our work and Ref.~\cite{Carullo:2021dui}.
In particular, as we will see in Sec.~\ref{sec:results}, adding theory-specific
information to the ParSpec coefficients \emph{can lead to different interpretations
of the bounds on $\ell_{\rm th}$, even for different theories that predict the same
value of the exponent $p$.}

@@

% --------------------------
\section*{Acknowledgements}
\label{sec:acknowledgements}
%
We thank Emanuele Berti, Andrea Maselli, Caio F. B. Macedo, Deyan Mihaylov,
Serguei Ossokine, Scott E. Perkins, and Helvi Witek for discussions.
%
We are grateful for the computational resources provided by the Max Planck
Institute for Gravitational Physics in Potsdam, specifically the
high-performace computing cluster Hypatia and to Steffen Grunewald for
assistance.
%
The material presented in this manuscript is based upon work supported by NSF’s LIGO Laboratory,
which is a major facility fully funded by the NSF.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
% General:
\redsout{The detection of gravitational waves produced by coalescing compact binaries
allows}
\hs{Detections of gravitational waves from coalescing compact binaries allow}
us to explore \hs{into} the dynamical, nonlinear regime of general relativity
and \redsout{to} constrain modifications to it.
%
Some of these gravitational-wave events have sufficiently high
signal-to-noise ratio in the merger, allowing us to probe the relaxation of the remnant black
hole to its final, stationary state --- the so-called black-hole ringdown, which is
characterized by a set of quasinormal modes.
% Main question:
Can we use the ringdown to constrain deviations \redsout{to} \hs{from} general relativity,
as predicted by several of its contenders?
% What we do:
Here, we  address this question by using an inspiral-merger-ringdown waveform
model in the effective-one-body formalism, augmented with a parametrization of
the ringdown based on an expansion in the final black hole's spin.
%
We give a prescription on how to include in this waveform model\hs{,} the quasinormal mode frequencies
calculated on a theory-by-theory basis. In particular, we focus on theories that modify
general relativity by higher-order curvature corrections, namely,
Einstein-dilaton-Gauss-Bonnet\hs{,} \redsout{and} dynamical Chern-Simons theories, and cubic- and quartic-order
effective-field-theories of general relativity.
%
% Main conclusion:
\redsout{We confront the resultant waveform models against the two loudest}
\hs{We use this parametrized waveform model to measure the ringdown properties of}
the two loudests \redsout{gravitational-wave} ringdown signals observed so far, \redsout{notably,} GW150914 and GW200129.
%
We find that while the Einstein-dilaton-Gauss-Bonnet theory cannot be
constrained with these events, we can place upper bounds on the
fundamental length-scale of cubic- ($\ell_{\rm cEFT} \leqslant 38.2$~km) and quartic-order
($\ell_{\rm qEFT} \leqslant 51.3$~km) effective-field-theories of general
relativity, and of dynamical Chern-Simons gravity ($\ell_{\rm dCS} \leqslant 38.7$~km).
%
The latter result is a concrete example of a theory presently unconstrained by
inspiral-only analyses which, however, can be constrained by merger-ringdown
studies with current gravitational-wave data.
%
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%

@@

In the end, the $\pSEOB$ waveform model with theory-specific
QNMs has only $\ell_{\rm th}$ as a free beyond-GR parameter.
%
We emphasize that our procedure is different from that of
Ref.~\cite{Carullo:2021dui} which, for a given value of $p$, varied all $\ell_{\rm th}$,
$\delta\omega^{(j)}$, and $\delta\tau^{(j)}$ parameters, and then proceeded
to use the posteriors on $\ell_{\rm th}$
\hs{, considering up to $j=2$ in
the GR deformation coefficients, and remaining agnostic about the underlying
theory which would predict the modifications to the QNMs.}
%
\redsout{to draw conclusions to any theory that predict\hs{s}
corrections at a given $p$.}
%
\redsout{This is a important difference between our work and Ref.~\cite{Carullo:2021dui}.
In particular, as we will see in Sec.~\ref{sec:results},}
\hs{We will see in Sec.~\ref{sec:results} that} adding theory-specific
information to the ParSpec coefficients \emph{can lead to different interpretations
of the bounds on $\ell_{\rm th}$, even for different theories that predict the same
value of the exponent $p$.}

@@

% --------------------------
\section*{Acknowledgements}
\label{sec:acknowledgements}
%
We thank Emanuele Berti, Andrea Maselli, Caio F. B. Macedo, Deyan Mihaylov,
Serguei Ossokine, Scott E. Perkins, and Helvi Witek for discussions.
%
\hs{We also thank Anuradha Gupta and Gregorio Carullo for comments on this work.}
%
We are grateful for the computational resources provided by the Max Planck
Institute for Gravitational Physics in Potsdam, specifically the
high-performace computing cluster Hypatia and to Steffen Grunewald for
assistance.
%
The material presented in this manuscript is based upon work supported by NSF’s LIGO Laboratory,
which is a major facility fully funded by the NSF.
